---
title: "Lab 11 BIMM 143"
format: pdf
editor: visual
---

# Section 1. Proportion of G/G in a Population

Downloaded a CSV file from Ensembl: <https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel>

Here we read this CSV file:
```{r}
mxl <- read.csv("mxl.csv")
head(mxl)
```
We see from the analysis below that 14.0625% of the Mexican Ancestry in Los Angeles are homozygous for the asthma associated SNP (G|G) 
```{r}
table(mxl$Genotype..forward.strand.)/ nrow(mxl) * 100
```
## Section 4. Population Scale Analysis

First, we read the text file we were given into a data frame. Then we look at the first couple rows using the `head()` function.
```{r}
expr <- read.table("expr.txt")
head(expr)
```
We check the number of rows to see how many observations we have in our dataset.
```{r}
nrow(expr)
```
We look at summarized version of the data we get the following:
```{r}
table(expr$geno)
summary(expr$geno)
```
Now we create a box plot representing the data in our table:
```{r}
library(ggplot2)
```

```{r}
ggplot(expr, aes(geno, exp, fill = geno)) + geom_boxplot(notch = TRUE)
```

